Cover letter for the original IJRR submission

### 2017-10-10 reviews_tro/reviewer_response.tex

\section{Summary of the modifications}

First of all, we would like to thank all the Reviewers and the associate editor (AE).
We received three detailed and technical reviews that helped improving our paper.
Overall we agreed with most of the reviews and took them into account in the revised version of the paper.
The main points of concern raised by the Reviewers regarded:
\begin{itemize}
\item the diffficulty to understand the takeaway message of the paper, and what was novel with respect to our previous paper [20] (and [1] to a lesser extent);
\item the organization of the paper, with too much information relegated to the Appendix, or only referred to in other contributions.
\end{itemize}

\textbf{Regarding the contribution of this paper with respect to our previous papers}, this paper must be understood
as an extension of [20] (presented at the ISRR conference). As a self-contained journal paper, the present paper aims at replacing [20]. We present two contributions. The first one, as in [20], is an efficient and generic planning method for multi-contact locomotion (problems P1 and P2).
The second one is a rigorous experimental validation of the approach on actual robot models, either in simulation or on the real HRP-2 robot.

Regarding the first contribution, the algorithm presented here is the same as in [20], with three novelties: a novel criterion for
static equilibrium, the pseudo-code of the algorithm,
as well as  the release of the source code of the planner.

@@

\textbf{Regarding the organization of the paper}, in the initial version we tried to synthesize as much as possible
the core of the paper, while providing the most technical details in the appendix. Based on the Reviewers comments, we modified
the paper architecture in such a way that all the novel details are now directly included in the core of the paper. This should avoid the reader going back and forth
in the paper, and also help him distinguishing between what is novel and what is not. As a result,  Appendices B.B (equilibrium criterion), D (pseudo code) and E (reference to the source code) are now integrated in the core of the paper. The manipulability-based heuristics used for contact generation, as well as our solution to the problem P3 remain in the Appendix as they are not novel. However, as suggested by the Reviewers, they are described in more details in the core.

Following the recommendation of the associate editor who ruled over some contradictory opinions of the Reviewers, we made the paper self-contained,
in such a way that the paper [20] is now obsolete with respect to this submission, and is only mentioned in the introduction, or to emphasize an improvement
of the original algorithm. The missing information pointed out by the Reviewers has been 
included in the paper, as explained in the detailed answers that follow. As a result, we believe that the new version of the paper is self-contained and can be understood if read linearly.\\

@@

 %~ In order to try to make the paper more readable, the first version was mixing three styles: details given in the core of the paper; details given in the appendices; details deferred in cited papers. While all Reviewers agreed that using all three styles together was a bad idea, we received contradictory reviews about how to modify the paper. Following the recommandation of the AE, we have worked to make the paper self-contained with all the contributions in the core of the paper. 
%~ We have brought all the material from [20] in the revised paper so that no references to this preliminary paper is necessary anymore in the technical description, and removed trivial reference to [20] but in the introduction.
%~ We merged Appendices B.B (equilibrium criterion), D (pseudo code) and E (referernce to the source code) in the core of the paper. We kept in the appendices the manipulability computations which is necessary for completeness but is well-known. We also prefered to keep the trajectory solver P3 in the appendix, although some details are novel, in order to keep clear that the main contributions of the paper are on P1 and P2.

We have thus addressed the majority of the comments and questions raised by the Reviewers. In our opinion, these modifications result in a significant improvement of the paper.

@@

\quot{
(4)	The last contribution is validation using an in-house
simulator, as opposed to more approximate (unrealistic avatar) models
used in previous work.	As with the previous contributions claimed, you
should flesh out just a bit more detail, along with a roadmap to where
a particular contribution appears in the paper, to prepare and guide
the reader a bit more. More precisely, what are the differences between
these two simulation approaches (previous and current)? You need a
better description of what your in-house simulator does and does not
do; a vague citation to another paper based on that simulator [38] is
not sufficient.}

\as{Thanks for the comment. Regarding the absence of details on the dynamic simulator, first of all we apologize because [38] is not the appropriate reference.
We added a section VIII.A where we describe the simulation in more details.

Regarding the difference between our previous and current work, the postures computed in [20] were not validated in a simulator, nor the complete motions since 
we did not have a solution to P3 at the time. Then the main
difference between the avatars of [20] and the real robots we now address are the joint limitations and auto-collisions probabilities.
The robots are much more constrained in this regard, which makes the planning harder. We thus had to demonstrate that our algorithm is also working
for these models, while maintaining good performance.
We added this missing piece of information to the paper, at the end of Section I.B.}
\qt{(Section VIII.A) To generate continuous movements from our contact plans we used either the framework proposed in [1], or our own implementation of a $\mathcal{P}_3$ solver (Appendix C).
The resulting movements have been validated either on the real HRP-2 robot (details can be found in [1]), or with our dynamic simulator, based on a state-of-the-art algorithm [25].
In the simulations we controlled the robot with a standard inverse-dynamics controller[31]. The code source of the simulator is available at \url{https://github.com/andreadelprete/pinocchio_inv_dyn/releases/tag/rbprm}.
This controller tries to follow the given whole-body trajectories, giving higher priority to the center-of-mass and end-effectors tracking with respect to the joint tracking.
The controller also makes sure that the resulting contact forces lie inside the specified friction cones (we used a friction coefficient of 0.3), and that the joint position, velocity and torque limits are satisfied.
The companion video shows the obtained motions.}
\done
%~ \textcolor{red}{TODO ANDREA}.

\quot{p.14 - not a good description of the simulation software used [\dots]Ref [38] focuses on robustness of simulations, but presumably, the testing in the present work is deterministic.
 just using the same simulator? Not much description is given of how/whether contacts/slipping are modeled [\dots] Why isn't a 3rd party software used, or at least software with better documentation ? Better info is needed to understand what the simulation truly represents.}
\as{Thanks for the comment. In summary, we had to implement our simulator because we were not satisfied with commercial simulators. Our simulator implements the paper ``Staggered projections for frictional contact in multibody systems'', by Kauffman, Sueda, James and Pai, which corresponds to a state-of-the-art solution to the problem.
We considered other simulators, namely Mujoco, ODE, Gazebo, Bullet and Open-HRP. We discarded Mujoco because we work on Linux and it only provides a windows implementation. In our tests, we were not able to obtain coherent results with ODE and Gazebo, even for simple pendulum models. At the time of choosing which simulator to use, Bullet did not allow to express the dynamics
equations using joint coordinates, but only cartesian coordinates, which made the formulation much more complex. Finally, Open-HRP is a great tool, but it only provides an interface for position-controlled models, and not torque-controlled. For these reasons, and other historical ones, we implemented our own simulator. We understand that because we use an in-house implementation, it is harder to trust the results we obtain. However, the implementation is from a state of the art LCP solver, used by many in the community. In this new submission, we also provide the source of the simulator. As for the paper itself and the rest of the code, we put a considerable effort in providing the reader with all the elements that will lead to an objective opinion of the paper.}

@@

\quot{What about the (unavoidable) uncertainty if a world model acquired via
on-line sensing is considered ? Perhaps some of these issues (e.g.
uncertainty) are not in the scope of the paper, but I can imagine that
a robotic researcher who reads the paper and is about to use the open
source software it advertises is very interested in the answers.}
\as{That is a good observation. Although the integration of on-line sensing and state estimation in the planner is an issue currently considered by our team, it is indeed
out of the scope of this paper. As reported in the previous answer we indicated it in Section III. We think that our planner could work similarly with 
a sensing system that would output data easily interpreted as a polygon soup (as for a LIDAR), otherwise additional treatment would be required. In any case additional work would be required to plan motions robust to the uncertainty of the sensing and the estimation. } \done

\quot{I appreciate the thorough presentation of results in the paper, however
it seems that this presentation is somewhat biased towards simulation
results. Simulations are perfectly OK for planning algorithms, but if
you decided to present the experimental results (HRP-2), then at least some
illustrations should be included in the paper. now they are only in the
accompanying video.}
\as{We agree with the remark, and included some illustrations of the experimental results on HRP-2 in the paper.}
%~ We now elaborate on the validity of the simulation with respect to an experiment on the real robot.

\done

 \quot{also a comparison between the results of the same
task(s) executed in dynamic simulation and on the real robot should
convince the reader that the planning method's good performance does
not depend too much on the idealized simulation environment. }
\as{As also suggested by Reviewer 1, we added a paragraph
in the paper (Section VIII.A)  to provide more information on the simulation we use to validate our motions. This simulation is an implementation of the state of the art approaches used by the community. Of course, it does not guarantee that the motion could be executed on the robot, but it is used extensively by the community as a reasonable feasibility test. The suggestion of the Reviewer to compare the simulation results with those on the real robot is interesting. However, in our simulations HRP-2 is torque controlled, whereas in our experiments HRP-2 is position controlled. This makes the usefulness of a comparison between simulation and experiments extremely doubtful, because we would be comparing motions obtained with different controllers. For these reasons, we prefer not to do this comparison, although we plan to do it in the future, since we are currently working on making HRP-2 torque controlled.}
\qt{(Section VIII.A) To generate continuous movements from our contact plans we used either the framework proposed in [1], or our own implementation of a $\mathcal{P}_3$ solver (Appendix C).
The resulting movements have been validated either on the real HRP-2 robot (details can be found in [1]), or with our dynamic simulator, based on a state-of-the-art algorithm [25].
In the simulations we controlled the robot with a standard inverse-dynamics controller[31]. The code source of the simulator is available at \url{https://github.com/andreadelprete/pinocchio_inv_dyn/releases/tag/rbprm}.
This controller tries to follow the given whole-body trajectories, giving higher priority to the center-of-mass and end-effectors tracking with respect to the joint tracking.
The controller also makes sure that the resulting contact forces lie inside the specified friction cones (we used a friction coefficient of 0.3), and that the joint position, velocity and torque limits are satisfied.
The companion video shows the obtained motions.}

### 2017-10-10 body/intro.tex
nicolas comments

\noindent \textbf{Regarding $\mathcal{P}_1$}. We propose a fast guide path planning algorithm. The key to its efficiency is that it does not sample directly the contact manifold, but an approximation of the \textit{contact reachable} space. The \textit{contact reachable} space is a low-dimensional space for which there exists a mapping to the contact manifold.

@@

The reachability condition is the key to the strict separation between $\mathcal{P}_1$ and $\mathcal{P}_2$, hence to the low complexity of our planner, though it
can result in failures. However, we demonstrate empirically its interest: the high success rate and low computation times allow us to plan (and re-plan upon failure) multi-contact sequences at \gls{interactive} rates.

@@

The other contribution of the paper is a rigorous experimental validation of the approach on actual robot models (Section~\ref{sec:results}).
To validate our contact plans we introduced a complete framework for multi-contact motion synthesis. This framework additionally comprises an interpolation method to solve the problem $\mathcal{P}_3$,  based on a reformulation of our previous work~\cite{Carpentier2016}. Our solution to $\mathcal{P}_3$ allows us to verify that the synthesized motions are physically consistent, using our implementation of a state-of-the-art simulation algorithm~\cite{Kaufman2008}. 
We do not consider these additions as contributions \textit{per se}, but describing them here is required to understand our experimental contribution.

Finally, as opposed to our previous paper, we demonstrate the algorithms with actual robotic models. These models are more constrained kinematically than the virtual avatars used before,
thus making the planning problem harder. Through an extensive empirical analysis, we demonstrate that the method remains efficient and applicable in this context.
%~ These aspects of the framework are presented in details in the paper, but are not novel \textit{per se}. The novelty lies in the validation of the method with real robot models, for which the planning is harder with respect to the avatars used in [20], because of stricter kinematic constraints

>>>>

\noindent \textbf{Regarding $\mathcal{P}_1$}, we propose a fast guide path planning algorithm. The key to its efficiency is that it does not sample directly the contact manifold, but an approximation of the \textit{contact reachable} space. The \textit{contact reachable} space is a low-dimensional space for which there exists a mapping to the contact manifold.

@@

The reachability condition is the key to the strict separation between $\mathcal{P}_1$ and $\mathcal{P}_2$, hence to the low complexity of our planner. However the reduction
can result in failures. We demonstrate empirically its interest, through an extensive experimental validation with real robot models on a dynamic simulator. The high success rate and low computation times we obtain allow us to plan (and re-plan upon failure) multi-contact sequences at \gls{interactive} rates.

@@

The other novelty of the paper is a rigorous experimental validation of the approach on actual robot models (Section~\ref{sec:results}).
To validate our contact plans we introduced a complete framework for multi-contact motion synthesis. This framework additionally comprises an interpolation method to solve the problem $\mathcal{P}_3$,  based on a reformulation of our previous work~\cite{Carpentier2016}. Our solution to $\mathcal{P}_3$ allows us to verify that the synthesized motions are physically consistent, using our implementation of a state-of-the-art simulation algorithm~\cite{Kaufman2008}. 
These aspects of the framework are presented in details in the
paper, but are not novel \textit{per se}. %Describing them here is required to understand our experimental contribution.
The novelty lies in the complete
validation of the contact planner with real robot models, for
which the planning is much harder with respect to the avatars
used in~\cite{tonneauisrr15}, because of more restrictive kinematic constraints.

%~ Finally, as opposed to our previous paper, we demonstrate the algorithms with actual robotic models. These models are more constrained kinematically than the virtual avatars used before,
%~ thus making the planning problem harder. Through an extensive empirical analysis, we demonstrate that the method remains efficient and applicable in this context.
%~ These aspects of the framework are presented in details in the paper, but are not novel \textit{per se}. The novelty lies in the validation of the method with real robot models, for which the planning is harder with respect to the avatars used in [20], because of stricter kinematic constraints
